\section{Smartcard Access Extension to Ruby}\label{impl:rubygem}
I have implemented a Ruby extension that provides access to PC/SC smart card
readers, and was tested to work on Windows XP and Vista, Mac OS X Tiger and
Leopard, as well as Ubuntu Linux versions 7.10 and 8.04. The extension is
packaged using RubyGems \cite{berube2007forge}, the de-facto Ruby software
packaging standard. The extension has been placed under the MIT license, and
released on the Rubyforge site \cite{berube2007forge}, which makes the gem
readily avaible on any RubyGems-enabled Ruby interpreter, via the command
\texttt{gem install smartcard} .

\subsubsection{Motivation}
The first TEM driver was implemented in Java. The choice seemed natural, as my
JavaCard IDE came with Java libraries that interacted with the cards. However,
Java's lack of advanced language features makes it inadequate for assembling
SECpacks. Writing a proper assembler would have required a lot of effort, and
assembling SECpacks by hand became unattractive once I started using complex
SECpacks. This impacted research negatively, as I had less enthusiasm to
experiment.

The opportunity for tossing away the Java code presented itself when a
requirement was introduced that the TEM's demo software must be able to run on
both Mac OS X, and Linux. Java's libraries for smart card access were not well
developed for Mac OS X, so at that point it became reasonable to invest effort
into writing a Ruby extension, as opposed to a Java JNI library.

\subsubsection{Features}
When running under Windows, the gem uses the native PC/SC access library,
\texttt{winscard}. Under Linux or Mac OS X, the gem uses the \texttt{pcsclite}
library, which is a free port of a subset of \texttt{winscard}. All the
functions in \texttt{pcsclite} are exposed using an object-oriented interface,
and the gem code works around the peculiarities\footnote{bugs} in the PC/SC
implementations it targets.
