\section*{Acknowledgments}

I would like to dedicate this work to my advisor, Prof. Srinivas Devadas. Words
cannot express my gratitude for all the trust and learning oportunities he has
provided.

My design would have been significantly less elegant had I not encountered Ruby
on Rails. To that end, I must thank Yukihiro Matsumoto (a.k.a. Matz) for
creating Ruby, as well as David Heinemeier Hansson for enabling the language to
reach the public spotlight by building the Rails framework.

I owe the thoroughness of my design to Luis Sarmenta. The TEM architecture
would have had a lot more rough edges had it not been for his endless
willingness to listen to my ideas, ask clarifying questions, and give feedback.

Marten van Dijk's words pushed me to come up with a design that is both elegant
and flexible, so that mobile agents can be easily implemented on the TEM. Marten
also had the patience to review my thesis, and his advice on real life issues was
invaluable.

I owe special thanks to Richard Kilmer, Jim Weirich, Chad Fowler, and Eric
Hodel. Their packaging system, rubygems, saved me countless hours. Also, the
rubyforge.org website, maintained by them, provided me with a very easy way to
provide my code to those I collaborated with.

The OpenSSL integration in my TEM implementation is owed to Jacob Strauss, who
wrote the custom OpenSSL cryptographic engine, as well as provided me with
guidance and testing.

I gained significant design skills and exposure to systems research thanks to
Professors Robert Morris and Franz Kaashoek and their wonderfully inspiring
courses Operating Systems Engineering (MIT course number 6.828) and Distributed
Systems. (MIT course number 6.824)

The names Yue and Mi were inspired by the movie ``Rush Hour 3'',
as well as by my two students with strikingly similar names, Yue Yang (Cherrie)
and Yueyang Li (Alice). Special acknowledgements go to Chengxi Huang (Cathy).
I wouldn't have even noticed the students' names if she wouldn't have given
me \texttt{teh y3110w 43v3r}.
